---
title: "INFO201 BB10 E-Commerce Price/Product Analysis"
author: "Gabe Kan, Shin Jingjang, Yaya Zheng"
editor: visual
format:
  html:
    embed-resources: true
---

# Part I: Find the Data

```{r}
#| warning: false
library(dplyr)
library(tidyverse)

# Import Datasets
amazon_products <- read.csv("amazon.csv")
walmart_products <- read_csv("walmart.csv")

# Summarize data sets (Also Converted Rupees in Amazon dataset to USD)
# str_remove_all() removes all of that character from the string vector
amazon_summary_df <- amazon_products |>
  reframe(
    product_name,
    rating_count = rating_count |> as.numeric(),
    average_rating = rating,
    actual_price_usd = round(
      actual_price |>
        str_remove_all("₹") |>
        str_remove_all(",") |>
        as.numeric() * 0.012,
      2
    )
  )
walmart_summary_df <- walmart_products |>
  reframe(
    product_name = Title,
    rating_count = `Num Of Reviews` |> as.numeric(),
    actual_price_usd = Price,
    average_rating = `Average Rating`
  )

# Clean Amazon and Walmart Dataset Summary
walmart_summary_df <- walmart_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )

amazon_summary_df <- amazon_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )
```

## **Data Source**

Our datasets come from Kaggle:

-   **Amazon Sales Dataset**:<https://www.kaggle.com/datasets/karkavelrajaj/amazon-sales-dataset>

-   **Walmart Product Data**:<https://www.kaggle.com/datasets/promptcloud/walmart-product-data>

Both contain real product listings from Amazon and Walmart, including product names, prices, review counts, and average ratings.

## **Data Meaning**

The datasets represent online products from Amazon and Walmart. Key fields: product name/title, price, review count, and average rating. They allow comparisons of pricing and customer engagement across platforms.

## **Data Join**

To create comparable product pairs:

-   Converted Amazon prices to USD and numeric review counts

-   Removed missing product names, reviews, or ratings

-   Created price_range (Low-, Medium-, High-Range)

-   Ranked top 100 most-reviewed products per price tier\

We then performed a **left join** on price_range and rank, keeping only Amazon products with a Walmart match.

## **Data Observations**

-   Columns: product names, prices, review counts, average ratings, price range, rank

-   Each row represents a matched Amazon–Walmart product within the same price tier and review rank

## **Data Issues**

-   Missing product names, reviews, or ratings (filtered out)

-   Amazon prices required cleaning and conversion

-   Limited products in some price tiers reduced matched pairs

Cleaning ensured the joined dataset is reliable for analysis.

# Part II: Data Join

```{r}
#| warning: false
# Mutate and Modify Columns, Add Price Range and Sum up Total Reviews 
# case_when() works as an "if...else" scenario, on "dplyr" documentation

amazon_summary_df <- amazon_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

walmart_summary_df <- walmart_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

# Rank top 100 most reviewed per price tier in each dataset
# slice_max selects rows with the largest values of a variable
# (Side Note: The Amazon dataset doesn't have enough product data--specifically High-End and Low-End products-- to create a full top 100 (not even top 30 D:). Because of this, 
# the merge has to be a one-to-one with the Amazon product data, and we scrap the rest of the Walmart products (very inefficient).)
amazon_top100 <- amazon_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())

walmart_top100 <- walmart_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())


# Joined Datasets Together
combined_df <- amazon_top100 |>
  left_join(walmart_top100, join_by(price_range, rank)) |>
  rename(
    amazon_product  = product_name.x,
    amazon_price    = actual_price_usd.x,
    amazon_reviews  = rating_count.x,
    amazon_rating = average_rating.x,
    walmart_product = product_name.y,
    walmart_price   = actual_price_usd.y,
    walmart_reviews = rating_count.y,
    walmart_rating = average_rating.y
  )

glimpse(combined_df)

```

We combined product data from Amazon and Walmart to compare prices and review activity across platforms.

### **Data Preparation (High-Level)**

Before joining, we:

-   Converted Amazon prices from rupees to USD

-   Converted review counts to numeric values

-   Summarized the Amazon and Walmart datasets into columns: Product Name, Number of Reviews, Average Rating, Price, and Price Range.

-   Filtered out missing product names and review counts

-   Removed products with zero reviews

-   Created a price_range variable (Low-, Medium-, High-Range)

-   Ranked the top 100 most-reviewed products within each price range

-   Added a rank column per price tier

Instead of joining directly on product names, we matched products using **price_range and rank** to create comparable Amazon–Walmart pairs.

### **Join Method and Result**

We performed a left join from Amazon to Walmart on price_range and rank, then removed unmatched rows and renamed columns for clarity (e.g., amazon_price, walmart_price).

The final dataset contains:

-   Columns: Amazon and Walmart product names, prices, review counts, average rating, price_range, and rank

-   Rows: reported using glimpse(combined_df)

Each row represents a matched Amazon and Walmart product within the same price tier and popularity ranking, allowing direct comparison of pricing and customer engagement.

# Part III: Data Documentation

## **Dataset Sources and Codebook Information**

This project uses two publicly available datasets from Kaggle:

Walmart Product Data (PromptCloud):\
[https://www.kaggle.com/datasets/promptcloud/walmart-product-data](https://www.kaggle.com/datasets/promptcloud/walmart-product-data?utm_source=chatgpt.com)

Amazon Sales Dataset:\
[https://www.kaggle.com/datasets/karkavelrajaj/amazon-sales-dataset](https://www.kaggle.com/datasets/karkavelrajaj/amazon-sales-dataset?utm_source=chatgpt.com)

Neither dataset provides a formal or detailed official codebook. The Kaggle pages include high-level descriptions of the datasets but do not supply structured documentation defining each column in detail. Because these datasets were collected through web scraping, it is likely that a formal codebook was not provided by the original data source. Therefore, column meanings were inferred from variable names, dataset previews, and standard e-commerce data conventions (e.g., product title, rating count, price, and average rating).

## **Relevant Variables Used in the Group Project**

From the original datasets, the most relevant variables include:

Amazon dataset:

-   product_name

-   rating_count

-   rating

-   actual_price

Walmart dataset:

-   Title

-   Num Of Reviews

-   Price

-   Average Rating

During data cleaning and transformation, several new variables were created:

1.  actual_price_usd\
    Amazon prices were originally listed in Indian Rupees (₹). Currency symbols and commas were removed, values were converted to numeric form, and prices were converted to USD using a fixed exchange rate multiplier (0.012). Values were rounded to two decimal places to standardize price comparisons across platforms.

2.  price_range\
    A categorical variable created using case_when() to group products into:

-   Low-Range (under \$50)

-   Medium-Range (\$50–\$199.99)

-   High-Range (\$200 and above)

This allows structured comparison across similar pricing tiers.

3.  rank\
    Within each price_range, the top 100 products were selected based on rating_count using slice_max(). A ranking variable was created using row_number() to allow one-to-one comparison between Amazon and Walmart products within each price category.

4.  Prefixed variables after merging\
    After joining the datasets, variables were renamed to distinguish platforms:

-   amazon_product, amazon_price, amazon_reviews, amazon_rating

-   walmart_product, walmart_price, walmart_reviews, walmart_rating

The most meaningful variables for our data exploration are:

-   rating_count (product popularity)

-   average_rating (customer satisfaction)

-   actual_price_usd (standardized cost comparison)

-   price_range (tiered analysis structure)

-   rank (cross-platform matching mechanism)

These variables directly support our analysis of how product popularity and ratings compare across platforms within similar pricing tiers.

## **150-Word Statement: Potential Biases and Human Rights Principles**

Although the combined dataset enables structured comparison of product popularity and pricing, it presents limitations related to the eight human rights principles. Privacy is preserved because no personal user data is included. Transparency and explainability are supported through documented cleaning, currency conversion, ranking, and merging steps. However, accountability and professional responsibility require acknowledging assumptions such as fixed exchange rates and review-based ranking. The dataset lacks demographic and geographic variables, limiting evaluation of fairness and non-discrimination. It does not provide information about product safety or quality concerns, restricting assessment of safety and security. Because users cannot influence how products are ranked in this dataset, human control of technology is limited. While the project promotes open documentation and comparative analysis, it only partially supports the broader promotion of human values, as it does not address equity or consumer well-being directly.

The combined dataset makes comparing product popularity and pricing easier; it has some limitations related to the eight human rights principles. The dataset does not include any personal user information to protect their privacy. While the dataset is being clean we make sure to document all for currency conversion, ranking, and merging steps to ensure transparency and explainability.

# **Part IV: Combined Data Wrangling**

```{r}
#| warning: false
# Convert Walmart and Amazon Rating to Numeric, filter out NA rating values
combined_df <- combined_df |>
  mutate(
    amazon_rating = amazon_rating |> as.numeric(), 
    walmart_rating = walmart_rating |> as.numeric()
    ) |>
  filter(!is.na(amazon_rating), !is.na(walmart_rating))
# Add average total rating
combined_df <- combined_df |>
  group_by(price_range) |>
  mutate(
    amazon_avg_rating  = round(mean(amazon_rating,  na.rm = TRUE), 2),
    walmart_avg_rating = round(mean(walmart_rating, na.rm = TRUE), 2)
  )

```

Ultimately, we wanted to create an average rating for the products in each price range. To do this, we converted the value of the average rating for each product into a numerical value, so we can manipulate the variable and find the mean of each price range. Afterwards, we cleared out all NA values from both the Amazon and Walmart data sets, in case the conversions to numeric didn’t work for those specific products. Lastly, we created a new column that means the average rating for the products in each category. 

We thought this would be useful because examining the average rating for the products in each price range can tell us more about the specifics of how each price range contains products that users like, which can be helpful to know for product advertisement and other statistical analysis.

# Section V: Providing Group Information

TA: **Sophia Dai**

Section BB

Gabe Kan (CSE intended, University of Washington) [gkan010\@uw.edu](mailto:gkan010@uw.edu) is interested in how e-commerce has become an integral part of our daily lives and how to manipulate the data to show patterns in the ways users shop for products.

Shin Jingjang (Informatics/ statistics intended, University of Washington) [skira\@uw.edu](mailto:skira@uw.edu) is interested in competitiveness of e-commerce circle and would like to know how social impacts can impact of sale and reviews of different e-commerce company.

Yaya Zheng (Interaction Design, University of Washington) [xiaolan\@uw.edu](mailto:xiaolan@uw.edu) is interested in how product popularity and pricing could affect a customer’s purchasing decision.

# Section VI: Visualization



