---
title: "Product Analysis Dashboard"
author: "INFO 201, Group BB10"
format: 
  dashboard:
    theme: [cosmo]
    embed-resources: true
---
```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)

amazon_products <- read.csv("amazon.csv")
walmart_products <- read_csv("walmart.csv")

amazon_summary_df <- amazon_products |>
  reframe(
    product_name,
    rating_count = rating_count |> as.numeric(),
    average_rating = rating |> as.numeric(),
    actual_price_usd = round(
      actual_price |>
        str_remove_all("₹") |>
        str_remove_all(",") |>
        as.numeric() * 0.012,
      2
    ),
    brand = "Amazon"
  )

# Summarize data sets (Also Converted Rupees in Amazon dataset to USD)
# str_remove_all() removes all of that character from the string vector

walmart_summary_df <- walmart_products |>
  reframe(
    product_name = Title,
    rating_count = `Num Of Reviews` |> as.numeric(),
    actual_price_usd = Price,
    average_rating = `Average Rating` |> as.numeric(),
    brand = "Walmart"
  )

# Clean Amazon and Walmart Dataset Summary
walmart_summary_df <- walmart_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )

amazon_summary_df <- amazon_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )
# Mutate and Modify Columns, Add Price Range and Sum up Total Reviews 
# case_when() works as an "if...else" scenario, on "dplyr" documentation

amazon_summary_df <- amazon_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

walmart_summary_df <- walmart_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

# Rank top 100 most reviewed per price tier in each dataset
# slice_max selects rows with the largest values of a variable
# (Side Note: The Amazon dataset doesn't have enough product data--specifically High-End and Low-End products-- to create a full top 100 (not even top 30 D:). Because of this, 
# the merge has to be a one-to-one with the Amazon product data, and we scrap the rest of the Walmart products (very inefficient).)
amazon_top100 <- amazon_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())

walmart_top100 <- walmart_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())


# One-to-one match by price_range, and then rank, keeping only Amazon rows with a Walmart match
# rename() renames each variable, could have just renamed it from summary but I am lazy. 
combined_df <- amazon_top100 |>
  left_join(walmart_top100, join_by(price_range, rank)) |>
  rename(
    amazon_product  = product_name.x,
    amazon_price    = actual_price_usd.x,
    amazon_reviews  = rating_count.x,
    amazon_rating = average_rating.x,
    walmart_product = product_name.y,
    walmart_price   = actual_price_usd.y,
    walmart_reviews = rating_count.y,
    walmart_rating = average_rating.y
  )

# Convert Walmart and Amazon Rating to Numeric, filter out NA rating values
combined_df <- combined_df |>
  filter(!is.na(amazon_rating), !is.na(walmart_rating))
# Add average total rating
combined_df <- combined_df |>
  group_by(price_range) |>
  mutate(
    amazon_avg_rating  = round(mean(amazon_rating,  na.rm = TRUE), 2),
    walmart_avg_rating = round(mean(walmart_rating, na.rm = TRUE), 2)
  )

rating_summary <- combined_df |>
  group_by(price_range) |>
  summarise(
    amazon_avg = mean(amazon_rating, na.rm = TRUE),
    walmart_avg = mean(walmart_rating, na.rm = TRUE)
  )
overview_df <- combined_df |>
  ungroup() |>
  summarise(
    avg_amazon_rating = mean(amazon_rating, na.rm = TRUE),
    avg_walmart_rating = mean(walmart_rating, na.rm = TRUE),
    amazon_total_reviews = sum(amazon_reviews),
    walmart_total_reviews = sum(walmart_reviews)
  )

# New merged dataset for comparitive graph, <
both_df <- bind_rows(amazon_top100, walmart_top100)
```
# Introduction

## Row

:::{.card title="Project Overview"}

Online retail platforms like Amazon and Walmart play an important role in modern purchasing behavior. Product ratings and reviews can influence how consumers perceive and evaluate each product. Pricing tiers also help shape how users perceive the value of a product. Knowing how these elements work together improves our understanding of how online marketplaces function.

**Question 1:** Within the same price range and popularity rank, does Amazon or Walmart tend to have higher average product ratings?

**Question 2:** Do Amazon and Walmart have different average ratings across price ranges?

**Question 3:** Is there a correlation between product price and customer ratings across platforms?

To explore these questions, we used two product listing datasets collected from Amazon and Walmart. These datasets contain product titles, prices, review counts, and average ratings. We cleaned and standardized the data before joining them to create a matched comparison dataset. We chose to analyze Amazon and Walmart because they are two of the largest e-commerce companies in the world, controlling a significant portion of the global market share. Studying these companies allows us to better understand customer behavior in online shopping.

:::

# Overview
Before going through the summary statistics for our distribution and comparisons, these
are the key metrics for our top 100 products in each price range for Amazon and Walmart.

average rating: helps us understand and compare overall customer satisfaction on each platform.
total number of reviews: indicates customer engagement and product popularity.

These metrics provide a quick snapshot of the overall performance of Amazon and Walmart.
## Column {height=15%}
```{r}
#| content: valuebox
#| title: "Amazon Avg Rating"
list(
  icon  = "star-fill",
  color = "#FF9900",
  value = round(pull(overview_df, var = avg_amazon_rating), 2)
)
```

```{r}
#| content: valuebox
#| title: "Walmart Avg Rating"
list(
  icon  = "star-fill",
  color = "#0071CE",
  value = round(pull(overview_df, var = avg_walmart_rating), 2)
)
```

```{r}
#| content: valuebox
#| title: "Amazon Total Reviews"
list(
  icon  = "person-fill",
  color = "#FF9900",
  value = pull(overview_df, var = amazon_total_reviews)
)
```

```{r}
#| content: valuebox
#| title: "Walmart Total Reviews"
list(
  icon  = "person-fill",
  color = "#0071CE",
  value = pull(overview_df, var = walmart_total_reviews)
)
```
## Row {height=55%}
Next, we examine the distribution of product ratings across both platforms. This histogram will show how ratings are spread out among the top 100 products and highlight any differences in the customer feedback patterns between Amazon and Walmart.
```{r}
#| title: "Rating Distribution"
ggplot(both_df, aes(x=average_rating, fill=brand)) +
  geom_histogram( color="#e9ecef", binwidth = 0.25, linewidth = 0.3, position = 'dodge') +
  scale_fill_manual(values=c("#FF9900", "#0071CE")) +
  scale_x_continuous(breaks = 1:5) + 
  labs(x = "Star Rating", y = "Number of Products", fill = "") +
  theme_minimal(base_size = 13)
```
:::{.card title="About This Chart"}
The histogram shows the distribution of product ratings on Amazon and Walmart. Most products on both platforms receive ratings between 3.5 and 5 stars, suggesting generally positive customer feedback. Amazon appears to have a slightly higher concentration of products with ratings above 4 stars compared to Walmart. It is worth it to mention that this data comes from only the top 100 items of each price range, sorted by average rating. In the raw dataset containing even items with only one total review, Walmart has a larger spread across the average ratings.
:::
# Amazon/Walmart Analysis

## Row {height=45%}

### Column {width=60%}
```{r}
#| title: "Average Ratings by Price Range"
rating_by_price <- both_df %>%
  group_by(brand, price_range) %>%
  summarise(avg_rating = mean(average_rating, na.rm = TRUE), .groups = "drop")

ggplot(rating_by_price,
       aes(x = price_range, y = avg_rating, fill = brand)) +
  geom_col(position = "dodge") +
  labs(x = "Price Range", y = "Average Rating", fill = "Platform") +
  scale_fill_manual(values = c("Amazon" = "#FF9900", "Walmart" = "#0071CE")) +
  theme_minimal()
```

### Column {width=40%}

:::{.card title="About This Chart"}
This bar chart compares the average product ratings across different price ranges for Amazon and Walmart. The visualization helps identify whether higher-priced products tend to receive better ratings. In general, ratings remain relatively consistent across price ranges, suggesting that product price does not strongly influence customer satisfaction.
:::

## Row {height=45%}

### Column {width=60%}
```{r}
#| title: "Relationship Between Price and Rating"
ggplot(both_df,
       aes(x = actual_price_usd, y = average_rating, color = brand)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Product Price (USD)", y = "Average Rating", color = "Platform") +
  scale_color_manual(values = c("Amazon" = "#FF9900", "Walmart" = "#0071CE")) +
  theme_minimal()
```

### Column {width=40%}

:::{.card title="About This Chart"}
This scatter plot examines the relationship between product price and average rating. Each point represents a product, and the trend line shows the overall pattern for each platform. The plot suggests that there is little strong correlation between price and rating, meaning more expensive products do not necessarily receive higher customer ratings.
:::
# Data Context
## Column{height = 60%}
:::{.card title="Dataset 1: Amazon Sales Dataset"}

https://www.kaggle.com/datasets/karkavelrajaj/amazon-sales-dataset 

The Amazon dataset was posted by a Kaggle user, Karkacelraja J. This dataset contains product listings scraped from the online marketplace in January 2023. The dataset includes more than 1,000 product listings where each observation represents a single product listing, which includes details such as the product name, price, number of ratings, and overall customer rating. With over a thousand observations in the dataset, we can examine trends in consumer involvement and product popularity. Product prices in the dataset were originally listed in Indian Rupees, which required cleaning and conversion into U.S. dollars for comparison with the Walmart dataset.

One key benefit of this dataset is that it contains real world product listings which could be used to analyze patterns in product popularity and ratings. However, this dataset does not represent the full population of Amazon products as it was collected through web scraping and showing only parts of it. Some products and regions might also be represented differently depending on the web scraping process.
Amazon Dataset Cleaning
The raw Amazon dataset has many miscellaneous columns. We narrow it down to product name, rating, review count, and price. Prices are listed in Indian Rupees (₹) and are converted to USD by multiplying by 0.012.


:::


:::{.card title="Dataset 2: Walmart Product Data (PromptCloud)"}
https://www.kaggle.com/datasets/promptcloud/walmart-product-data

The Walmart dataset contains product listings scraped from the online marketplace. The dataset includes approximately 30,000 product records, each observation in the dataset represents a single product and includes variables like product title, price, number of reviews, and average rating. These could also allow us to analyze customer engagement and product popularity.

This dataset was compiled automatically by web scraping tools developed by PromptCloud. By using this tool, the information that actually reflects what is publicly visible on the website was created. 

One major benefit of this dataset is that it includes large numbers of product listings which give a broader view of Walmart’s online product. However, similarly to Amazon’s dataset, this dataset only represents parts of Walmart’s online product catalog and the review counts and ratings are changing over time.

Walmart Dataset Cleaning
The Walmart dataset has many columns with mostly missing values (carrier, screen size, etc.). We keep only product title, price, review count, and average rating, and filter out rows missing those key fields.

:::


## Row {height=40%}

:::{.card title="Data Relevance"}

These datasets include price, rating, and review data needed to study how pricing relates to customer satisfaction and popularity. Because both platforms provide the same measures, we can directly compare Amazon and Walmart to see whether price tiers and platform differences affect ratings.

:::

:::{.card title="Key fields"}
product name/title, price, review count, and average rating. They allow comparisons of pricing and customer engagement across platforms.

:::


```{r}
amazon_summary_df <- amazon_products |>
  reframe(
    product_name,
    rating_count = rating_count |> as.numeric(),
    average_rating = rating,
    actual_price_usd = round(
      actual_price |>
        str_remove_all("₹") |>
        str_remove_all(",") |>
        as.numeric() * 0.012,
      2
    )
  )
```


```{r}
# Summarize data sets (Also Converted Rupees in Amazon dataset to USD)
# str_remove_all() removes all of that character from the string vector

walmart_summary_df <- walmart_products |>
  reframe(
    product_name = Title,
    rating_count = `Num Of Reviews` |> as.numeric(),
    actual_price_usd = Price,
    average_rating = `Average Rating`
  )

# Clean Amazon and Walmart Dataset Summary
walmart_summary_df <- walmart_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )

amazon_summary_df <- amazon_summary_df |>
  filter(!is.na(rating_count),
         !is.na(product_name),
         !is.na(average_rating),
         rating_count > 0
  )
# Mutate and Modify Columns, Add Price Range and Sum up Total Reviews 
# case_when() works as an "if...else" scenario, on "dplyr" documentation

amazon_summary_df <- amazon_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

walmart_summary_df <- walmart_summary_df |>
  mutate(
    price_range = case_when(
      actual_price_usd < 50  ~ "Low-Range",
      actual_price_usd < 200 ~ "Medium-Range",
      actual_price_usd >= 200 ~ "High-Range"
    )
  )

# Rank top 100 most reviewed per price tier in each dataset
# slice_max selects rows with the largest values of a variable
# (Side Note: The Amazon dataset doesn't have enough product data--specifically High-End and Low-End products-- to create a full top 100 (not even top 30 D:). Because of this, 
# the merge has to be a one-to-one with the Amazon product data, and we scrap the rest of the Walmart products (very inefficient).)
amazon_top100 <- amazon_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())

walmart_top100 <- walmart_summary_df |>
  group_by(price_range) |>
  slice_max(order_by = rating_count, n = 100) |>
  mutate(rank = row_number())


# One-to-one match by price_range, and then rank, keeping only Amazon rows with a Walmart match
# rename() renames each variable, could have just renamed it from summary but I am lazy. 
combined_df <- amazon_top100 |>
  left_join(walmart_top100, join_by(price_range, rank)) |>
  rename(
    amazon_product  = product_name.x,
    amazon_price    = actual_price_usd.x,
    amazon_reviews  = rating_count.x,
    amazon_rating = average_rating.x,
    walmart_product = product_name.y,
    walmart_price   = actual_price_usd.y,
    walmart_reviews = rating_count.y,
    walmart_rating = average_rating.y
  )

# Convert Walmart and Amazon Rating to Numeric, filter out NA rating values
combined_df <- combined_df |>
  mutate(
    amazon_rating = amazon_rating |> as.numeric(), 
    walmart_rating = walmart_rating |> as.numeric()
    ) |>
  filter(!is.na(amazon_rating), !is.na(walmart_rating))
# Add average total rating
combined_df <- combined_df |>
  group_by(price_range) |>
  mutate(
    amazon_avg_rating  = round(mean(amazon_rating,  na.rm = TRUE), 2),
    walmart_avg_rating = round(mean(walmart_rating, na.rm = TRUE), 2)
  )

rating_summary <- combined_df |>
  group_by(price_range) |>
  summarise(
    amazon_avg = mean(amazon_rating, na.rm = TRUE),
    walmart_avg = mean(walmart_rating, na.rm = TRUE)
  )
avg_amazon_rating <- combined_df |>
  pull(amazon_rating) |>
  mean(na.rm = TRUE)
avg_walmart_rating <- combined_df |>
  pull(walmart_rating) |>
  mean(na.rm = TRUE)
```


# Group Information
## Row
:::{.card title="Group Information"}
TA: Sophia Dai

Section BB

Gabriel Kan (CSE intended, University of Washington) gkan010@uw.edu is interested in how e-commerce has become an integral part of our daily lives and how to manipulate the data to show patterns in the ways users shop for products.

Shin Jingjang (Informatics/ statistics intended, University of Washington) skira@uw.edu is interested in competitiveness of e-commerce circle and would like to know how social impacts can impact of sale and reviews of different e-commerce company.

Yaya Zheng (Interaction Design, University of Washington) xiaolan@uw.edu is interested in how product popularity and pricing could affect a customer’s purchasing decision.
:::

# Conclusion

## Row

:::{.card title="Project Conclusions"}

Our analysis revealed several patterns in how product pricing, ratings, and customer engagement appear across Amazon and Walmart. One major takeaway is that lower-priced products tend to accumulate a larger number of customer reviews. This pattern suggests that lower-priced products are likely purchased more frequently, which increases the opportunity for customers to leave reviews. As a result, these products often appear more popular when measured by review count.

Another finding from our analysis is that the overall average ratings across the two platforms are relatively similar. When comparing products within the same price ranges, both Amazon and Walmart show comparable rating distributions. This suggests that customers evaluate products in similar ways regardless of the platform where the product is listed. However, some small differences in ratings across price tiers may reflect differences in customer expectations for products at different price levels.

It is also important to consider the limitations of the datasets used in this project. Both datasets were collected using web scraping and only represent a subset of the full product catalogs from Amazon and Walmart. Because of this, our analysis may underestimate or overestimate certain trends in product ratings and review counts. Additionally, customer ratings can change over time as more reviews are added to products.

Future research could expand this analysis by including more complete datasets or by incorporating additional e-commerce platforms such as Target or eBay. Exploring more product categories or examining how review counts change over time could also provide deeper insights into how consumer behavior and online marketplaces interact.

:::